%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

%\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
 \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\usepackage{algorithm}
\usepackage{algorithmic}

\journal{Parallel Computing}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Improving Data Movement Performance for Sparse Data Patterns on the Blue Gene/Q Supercomputer}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[evl]{Huy Bui}
\author[mcs]{Eun-Sung Jung}
\author[mcs]{Venkatram Vishwanath}
\author[evl]{Andrew Johnson}
\author[evl]{Jason Leigh}
\author[alcf,niu]{Michael E. Papka}

\address[evl]{Electronic Visualization Laboratory (EVL), University of Illinois at Chicago, 842 W Taylor St, Chicago, IL 60607, USA}
\address[mcs]{Mathematics and Computer Science, Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, USA}
\address[alcf]{Argonne Leadership Computing Facility, Argonne National Laboratory, 9700 S Cass Ave, Lemont, IL 60439, IL, USA}
\address[niu]{Northern Illinois University, 300 Normal Road, DeKalb, IL 60115, USA}

\begin{abstract}

In-situ analysis has been proposed as a promising solution to glean faster insight and to reduce the amount of data written out to storage. A critical challenge here is that the reduced dataset needed to visualize a specific region of interest as the simulation is running is typically held on a subset of the nodes and needs to be written out to storage. Coupled multiphysics simulations also produce a sparse data pattern wherein data movement occurs among a subset of nodes. We evaluate the performance of these data patterns and propose several mechanisms for improving performance. Our mechanisms introduce intermediate nodes to implement multiple paths to transfer data on top of default routing algorithms and utilize topology-aware data aggregation to avoid shared bottleneck links. The efficacy of our solutions is evaluated through microbenchmarks and application benchmarks on an IBM Blue Gene/Q system scaling up to 131,072 compute cores. The results show that our algorithms achieve up to a 2X improvement in achievable throughput compared to the default mechanisms.

\end{abstract}

\begin{keyword}
multiple paths \sep sparse data movement \sep topology-aware aggregation \sep data-intensive \sep Blue Gene/Q
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%% \linenumbers

\input{intro}

\input{relatedwork}

\input{system}

\input{approach}

\input{microbenchmark}

\input{appbenchmark}

\input{conclusion}

\section*{Acknowledgments}

This research used resources of the Argonne Leadership Computing Facility (ALCF) at Argonne National Laboratory, and is supported by the Office of Science of the U.S. Department of Energy under contract DE-AC02-06CH11357. We thank the ALCF team for discussions and help related to this paper.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

\section*{References}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
\bibliographystyle{elsarticle-num} 
\bibliography{sigproc}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

%% \bibitem{}

%\end{thebibliography}
\end{document}
\endinput
%%
%% End of file `elsarticle-template-num.tex'.
