\section{Multi-path Data Movement/Routing}
\label{sec:approach}

In large-scale systems such as the Blue Gene/Q (BG/Q), data is routed through its interconnect using the default routing algorithms. They perform well for some communication patterns \cite{Chen:BGQ}. However, for certain communication patterns (shown later in this paper), the default routing algorithms result in poor performance due to unbalanced load on the physical network links. This results in a significantly larger amount of data being transferred over few links. On the BG/Q,  the default algorithm uses a single path to transfer data between any two nodes in the system. In addition, data traverses along fixed paths on certain links regardless of the overall load of the system. Thus, some links are overloaded while other links have less data or may even be idle. This overloading is a major bottleneck for data transfer throughput. Balancing the load while exploiting multiple physical network links can improve the throughput. 

The above problem can be formulated as a multi-commodity integral flow problem, which is shown to be NP-complete \cite{even1975}. Given a set of source and destination nodes, and the amount of data to be transferred from the sources to destinations, the problem is to find a set of paths from the source nodes to the destination nodes that results in high throughput. Additionally, the objective is to balance the overall system load in order to avoid congestion in the interconnect, and to avoid overloading physical network links. In this paper (Sections \ref{sec:heuristic} and \ref{sec:optimization}), we propose two approaches to solve this problem, taking into consideration the system topology. The first approach is a greedy heuristic that selects paths so that the load is balanced on the links. The second approach is an optimization based model that selects globally optimal paths for source-destination pairs. 

We leverage the idle or lightly-loaded links for data transfer in order to balance the load. For this we need to search for multiple paths between source and destination nodes and assign an appropriate load on each path. Present-day supercomputers have thousands of nodes and hundreds of thousands of edges due to complex interconnect topology. This implies a large search space for multiple paths. The brute-force approach of searching for paths can lead to significant amount of time being spent on searching load-balanced paths. We use Yen's algorithm \cite{Yen:Kpath} to search for a set of shortest paths. To reduce the search time we prune the search space by constraining the number of hops on each path. 

In order to search for paths, we model the interconnect network as a graph. Each compute node is modeled as a vertex and each physical link is modeled as an edge. Algorithm \ref{alg:h1} depicts the algorithm to generate $k$ shortest paths from a source to its destination, where $k$ is an input to the algorithm.
\begin{algorithm}[!htbp]
\SetKwFunction{FindKShortestPaths}{FindKShortestPaths}
\KwIn{Input:} Set of pairs of source-destination (\textit{s$_i$, d$_i$}). Graph of nodes. Desired number of shortest paths $k$.\\
\KwOut{Output:} Set of paths: $k$ paths for each source-destination pair.\\
\SetKwProg{myproc}{Procedure}{}{}
\myproc{\FindKShortestPaths{}}{
\For{ each pair of ($s_i$, $d_i$) }
{ 
  \While{(Less than $k$ paths have been discovered and there are more paths available)}
	{ 
 		Use Yen's algorithm to search for the shortest path $p$\;
		\If {Number of hops of p $\leq$ partition's network diameter}
		{
		    Add $p$ into $k\_paths$\;
		}
	}
}	
}	
\caption{$k$ shortest paths generation.}
\label{alg:h1}
\end{algorithm}

The input to the Algorithm \ref{alg:h1} includes a set of pairs of source-destination (\textit{s$_i$, d$_i$}), a graph of nodes and a desired number of shortest paths $k$ for each pair (\textit{s$_i$, d$_i$}). The output is a set of at most k paths for each pair of source and destination nodes. We search for paths by iterating through the set of pairs (line 3). For each pair, we search for one shortest path at a time using Yen's algorithm. If the found path has number of hops at most as the partition's network diameter, we add the path into the set of $k\_paths$\ of the pair (line 5-8). The search completes when we either have $k$ paths or there is no more paths found by Yen's algorithm (line 4). The paths returned by the Algorithm \ref{alg:h1} include both minimal and non-minimal paths. We can extend the search by allowing paths with number of hops higher than the partition's network diameter. However, it might lead to higher search time as well higher transfer time. In the next sections, we present our heuristic approach and optimization model that use $k\_paths$\ for finer searches for paths between pairs of sources and destinations.
