
The model is written in AMPL (A Mathematical Programming Language). The model is descrbied in \textbf{Model} 1.

\begin{algorithm}[!htbp]

\begingroup
\fontsize{9pt}{9pt}\selectfont

\begin{verbatim}

set Nodes;
set Arcs within Nodes cross Nodes;

set Jobs;
set Paths{Jobs};
set Path_Arcs{job in Jobs, p in Paths[job]} 
    within Arcs;

param Capacity{Arcs} >= 0 default Infinity;
param Demand {Jobs} default 0;

var Flow {job in Jobs, Paths[job]} >= 0;
var Z >= 0;

maximize obj: Z;

subject to

demand_con {job in Jobs}: sum {p in Paths[job]} 
	Flow[job,p] = Demand[job]*Z;

capacity_con {(i,j) in Arcs}:
  sum {job in Jobs, p in Paths[job]: 
    (i,j) in = Path_Arcs[job,p]} Flow[job,p] 
		<= Capacity[i,j];

\end{verbatim}

\endgroup

\caption*{\textbf{Model 1} Data movement optimization}
\label{mod:opt}

\end{algorithm}

The notions used in \textbf{Model} 1 are explained as follows:

\begin{itemize}
    \item sets: 
	\begin{itemize}
	    \item \textit{Nodes}: set of nodes in the network, each node represent a compute node in the supercomputer.
	    \item \textit{Arcs}: set of arcs in the network. Each arc represent a physical link in the supercomputer.
	    \item \textit{Jobs}: set of jobs. Each jobs has a source and a destination.
	    \item  \textit{Paths}: set of paths for each job.
	    \item \textit{Path\_Arcs}: set of arcs on each path of each job.
	\end{itemize}
    \item params: 
	\begin{itemize}
	    \item {\it Capacity}: capacity of each arc i.e. bandwidth of the physical link.
	    \item {\it Demand}: amount of data to be transferred of each job between a pair of source and destination.
	\end{itemize}
    \item vars:
	 \begin{itemize}
	    \item \textit{Flow}: flow of each job on a path. It can be seen as the proprotional bandwidth assigned for the job on that path.
	    \item \textit{Z}: is reversed of total time.
	\end{itemize}
    \item objective function: we want to minimize the time or maximize its reversed value i.e. maximize \textit{Z}.
    \item constraints(subject to): 
	\begin{itemize}
	    \item \textit{demand\_con}: flow of a job on equals to the demand of the job divided by the transfer time.
	    \item \textit{capacity\_con}: total flow on an arc is less than its capacity.
	\end{itemize}
\end{itemize}
