\subsubsection{Maxload value for heuristic approach}

For the heuristic approach, we use k shortest paths and a \textit{maxload} value to select the number of paths used for data transfer, as described in \ref{sec:heuristic}. Depending on the \textit{maxload} value, the heuristic may select a different set of paths which can affect performance. In this experiment, we show the effect of choosing the \textit{maxload} value and time to select paths based on the \textit{maxload} value. The experiment was carried out in a 1024-node partition, with 1 MPI/PAMI rank per node, 8 MB message size, source to destination ratio of 1:8 for all 3 patterns. The results for \textit{maxload} values of 1, 2, 4, 8, 16, and 32 are shown in Table \ref{table:maxload}. The first column shows the pattern type, the second column shows the MPI performance and the remaining columns show HEU performance for various \textit{maxload}.

\begin{table}[h]%[!htbp]
   \centering
    \begin{tabular}{| l |p{0.5cm} | p{0.5cm} |  p{0.5cm} | p{0.5cm} | p{0.5cm} | p{0.5cm} | p{0.75cm} |}
    \hline
     \multirow{2}{*}{Patterns} & \multirow{2}{*}{MPI} & \multicolumn{6}{ c| }{Maxload} \\ \cline{3-8}
     & & 1 & 2 & 4 & 8 & 16 & 32 \\ \hline
     Disjoint & 45 & 31 & 32 & 32 & 63 & 75 & 78 \\ \hline
     Overlap & 42 & 66 & 66 & 66 & 125 & 112 & 89 \\ \hline
     Subset & 74 & 69 & 70 & 69 & 114 & 110 & 96 \\ \hline
    \end{tabular}
    \caption{Throughput (GB/s) with different \textit{maxload} values for Heuristic approach.}
    \vspace{-0.15in}
    \label{table:maxload}
\end{table}

The performance of HEU with \textit{maxload} value of 1, 2 or 4 is similar, and lower than MPI\_Alltoallv in disjoint and subset pattern. This is because with lower \textit{maxload} values, the heuristic is not able to find enough paths to transfer data, leading to fewer number of physical links being used, thus higher load on those physical links. When the \textit{maxload} value is set to 32, performance starts to degrade because 
%in case of the Overlap and Subset patterns while only increasing slightly in the Disjoint pattern. 
%When the \textit{maxload} value set to a high value (32), 
the heuristic finds too many paths, leading to many paths sharing a physical link, thus leading to higher load on the physical links. The best performance is achieved with \textit{maxload} value of 8 and 16 because of better load distribution on the physical links. For the experiments in this paper we set \textit{maxload} value to 16.
% values we have an appropriate number of links, and the load is distributed better on the physical links. For the experiments in this paper we set \textit{maxload} value to 16.
\begin{comment}
When we increase the \textit{maxload} value, it also takes more time to select paths from the $k$ shortest paths. Table \ref{table:solvetime} shows the time for different \textit{maxload} values in different patterns.
\begin{table}[!htbp]
   \centering
   \begin{tabular}{| p {0.75cm}| r | r | r | r | r | r |}
    \hline
    \multirow{2}{*}{Pattern} & \multicolumn{6}{ c| }{Time for Different Max Load (s)} \\ \cline{2-7}
    & 1 & 2 & 4 & 8 & 16 & 32 \\ \hline
    Disjoint & 1.958 & 1.961 & 1.917 & 1.956 & 2.002 &  2.164 \\ \hline
    Overlap & 1.923 & 1.890 & 1.801 & 1.929 & 1.993 & 2.082 \\ \hline
    Subset & 1.907 & 1.870 & 1.891 & 1.955 & 2.024 &  2.223 \\ \hline
    \end{tabular}
    \caption{Search time with different max load in 1024 nodes partition.}
    \label{table:solvetime}
\end{table}

The search time is short and thus can be amortized over time when a pattern is used repeatedly.
\end{comment}
