\documentclass{acm_proc_article-sp}

\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\algrenewcommand\textproc{}
\usepackage{verbatim}

\begin{document}

\title{Improving Data Movement Performance by Balancing Network Load on the Blue Gene Q Supercomputer}

\numberofauthors{1}
\author{
% 1st. author
\alignauthor
Ben Trovato\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}

\maketitle
\begin{abstract}
Achievable networking performance in a supercomputer depends on the exact combination of communication patterns of applications and routing algorithms used by the supercomputer. However, while the communication patterns usually have wide spectrum and vary from application to applicaiton and phase to phase in an application, the routing algorithms have a limited variation and usually are optimized for typical communication patters. This results in high networking performance for favored communication patterns but low networking performance for other communication patterns. In this paper, we introduce approaches for improving networking performance by rebalancing load on physical links on the Blue Gene Q supercomputer. We realize our approaches into a framework called OPTIQ and demonstrate the efficacy of our framework via a set of benchmarks. The results show that we can achieve serveral times higher bandwidth than default mechannism MPI used in the Blue Gene Q supercomputer for certain communication patterns.
\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

\keywords{data movement, load balancing, optimization, communication patterns} % NOT required for Proceedings

\input{intro}

\input{relatedwork}

\input{system}

\input{approach}

\input{benchmark}

\input{conclusion}

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

\bibliographystyle{abbrv}
\bibliography{thesis}  % sigproc.bib is the name of the Bibliography in this case

\balancecolumns
% That's all folks!
\end{document}
